Typed Functional Genetic Programming

### 2013-07-14 tex/thesis.tex
[tex]

% !!! místo we can atd psát one can ...

@@

@@

\section{Generating individual}
\label{GPgene}

@@

In order to perform \textit{ramped half-and-half} method
one of those two sub-methods is randomly (uniformly) 
selected and the selected one is performed.\\

@@

A maximum depth $D_{initial}$ is defined, i.e. 6.

In nodes with \textit{depth} $\leq D_{initial}$ the \textit{full}
method selects from the set $F$, whereas the \textit{grow} method 
selects from whole \TuF.

In nodes with \textit{depth} $= D_{initial}$ both methods select from the
set $T$.

@@

\begin{todo}
   & určitě promluvit nějak v úvodu k tomu že se jedná o 
	 Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
	 řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
   & možná se hlobějc zmínit o curryingu  
\end{todo}

@@

An heuristic function takes as input a state and returns as output an estimation of number
of steps needed to get to a final state by the shortest possible path.

@@

Now we will describe further modification of the A* algorithm 
which is generalizing it enough to enable use of \textit{ramped half-and-half}
term generating method and other term generating methods.

\begin{enum}
	& Co se přidalo 
	  && strategy
	     &&& selectedNexts
	  && outer while loop 	
\end{enum}

@@

		$open    \la$ empty priority queue \;
		$strategy.init()$\;\;

@@

			\For{$next \in nexts$ }{
				$next$.G $\la $state$.G + 1$ \;
				$next$.F $\la $next$.G + heuristic$( $next$ )  \;
				$open$.insert( $next$ )\;			
			}
		}\;

	}

@@

\begin{todo}
 & Internal state
 & Edge selection model
\end{todo}

@@

>>>>

% !!! místo we can atd psát one can ...
% ujednotit i.e. , ty čaky kolem toho atd

% myslim že používam málo čárek, přečíst to a vygooglit kde mi příde že by mohla bejt

% ujednotit kurzívu a normální písmo v pseudocodu

@@

\begin{todo}
& koza tomu řiká closure, přečíst si tu pasáž (str 81) a 
upravit podle toho tady tu terminologii
\end{todo}

@@

\section{Generating individuals}
\label{GPgene}

@@

In order to perform \textit{ramped half-and-half} method
one of those two sub-methods is randomly (uniformly) 
selected and the selected one is performed.
This selection is done for each generated tree, so there
is 50\% chance to be generated by \textit{full} 
and 50\% chance to be generated by \textit{grow}.\\

A maximum depth $D_{initial}$ is defined (e.g. 6).

For each tree is selected its 
depth $d$ from $\setDots{2}{D_{initial}}$ randomly (uniformly).\\

@@

In nodes with \textit{depth} $\leq d$ the \textit{full}
method selects from the set $F$, whereas the \textit{grow} method 
selects from whole \TuF.

In nodes with \textit{depth} $= d$ both methods select from the
set $T$.

@@

\begin{todo}
   & určitě promluvit nějak v úvodu k tomu že se jedná o 
	 Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
	 řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
   & říct že se vždycky zastavěj (jak se to správně jmenuje?) ale
     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
     říct to někde kde už se ví jak sou napojený na termy    
   & možná se hlobějc zmínit o curryingu  
\end{todo}

@@

A heuristic function takes as input a state and returns as output an estimation of number
of steps needed to get to a final state by the shortest possible path.

@@

Now we will describe further modification of the A* algorithm 
which is generalizing it enough to enable use of \textit{ramped half-and-half}
term generating method and other term generating methods.\\

It is done by adding one more input argument for the algorithm.
This additional argument is a search strategy whose purpose is
to determine which successor states of all possible successor states 
will be added to the priority queue, i.e., a search strategy 
is there to filter out some successor states from $state.nexts()$.
This filtering may be non-deterministic. It is based on
the depth of expanded unfinished leaf and on properties of 
the newly added subtree, e.g., whether the newly added subtree
contains one or more unfinished leafs (so it corresponds to non-terminal) 
or not (so it corresponds to terminal).

By \textit{depth} we really mean $\lfloor depth/2 \rfloor$, because our 
formal state trees contain those type nodes above each symbol node. 
\\

In previous situation where there is no filtering of states involved 
we know that if queue becomes empty, then we have exhaustively searched
all possible states.

But since filtering is involved this is no longer true.
Therefore, if the queue becomes empty before the required number of
terms is generated, then the generating process is restarted and 
it continues to generate terms.

We also enable strategy to have some internal state which is 
initialized (possibly non-deterministically) each time the 
generating process restarts. \\

Hopefully those changes will be clarified by the following pseudocode:

@@

		$strategy.init()$\;\;

		$open    \la$ empty priority queue \;\;

@@

			\For{$next \in nexts$ }{
				$next$.G $\la $state$.G + 1$ \;
				$next$.F $\la $next$.G + heuristic$( $next$ )  \;
				$open$.insert( $next$ )\;			
			}
		}
	}\;

@@

\Lets show examples of search strategies. 
They are described by describing 
their $filter(nexts,depth)$ 
and $init()$ methods. Again, in order to avoid confusion, 
\lets stress that $depth$ is really $\lfloor depth/2 \rfloor$ since
our formal states contain above each symbol node a type node.

\subsubsection{Systematic strategy}

The previous systematic behavior may be obtained by trivial 
$filter$ function which returns all $nexts$
and $init$ is doing nothing.

\subsubsection{Ramped half-and-half}

In the $init$ are randomly (uniformly) initialized two variables of the
strategy internal state:

\begin{enum}
 & \textit{isFull} - A boolean value, determining whether \textit{full}
                     or \textit{grow} method will be performed.
 & \textit{d} - A integer value from $\setDots{2}{D_{init}}$, where 
                $D_{init}$ is predefined maximal depth (e.g. 6).                    
\end{enum}~

The method $filter(nexts,depth)$ returns precisely one randomly
(uniformly) selected  element from a subset of $nexts$
(or zero elements if $nexts$ is empty). 
This means that the queue always contains only one (or zero) state.

The subset to select from is determined by $depth, d$ and $isFull$.\\

Those elements of $nexts$ whose newly added subtree contains one ore more 
unfinished leafs are regarded as \textit{non-terminals}, whereas 
those whose newly added subtree contains no unfinished leaf are regarded as 
\textit{terminals}.\\

If $depth = 0$, then the subset to select from is  
set of all \textit{non-terminals} of $nexts$.

If $depth = d$, then the subset to select from is
set of all \textit{terminals} of $nexts$.\\

In other cases of $depth$ it depends on value of $isFull$.

If $isFull = true$, then the subset to select from is 
set of all \textit{non-terminals} of $nexts$.

If $isFull = false$, then the subset to select from is 
whole $nexts$ set.

\begin{todo}
& eště se musí nějak prohrotit to, že tam je navíc ta lambda hlava
kerá ještě o něco kazí tu $depth$
& ukázat že je to fakt to samý co kozovský klasický
\end{todo}

@@

& nekde zmínit že s těma nexts nezacházíme takle dementně, že samozřejmě vybíráme 
jen z těch přidávanejch podstromu ale že tim nechcem komplikovat popis
